RationalGRL: A Framework for Argumentation and Goal Modeling

### 2017-05-23 background.tex
Many many inconsistencies!

Most of the examples in this article, as well as the topic of discussion in the transcripts we analyze, come from the traffic simulator design exercise. In this exercise, designers are provided a problem description, requirements, and a description of the desired outcomes. The problem description is given in full in Appendix~\ref{ch:designprompt}, and is summarized as follows: The client of the project is Professor E, who teaches civil engineering at UCI. It is the task of the designer to specify a system in which the professor can teach students how various theories around traffic lights works, such as queuing theory. To this end, a piece of software has to be developed in which students can create visual maps of an area, regulate traffic, and so forth. The original version of the problem descrption~\cite{UCIworkshop} is well known in the field of design reasoning since it has been used in a workshop\footnote{\url{http://www.ics.uci.edu/design-workshop/}}, and transcripts of this workshop have been analyzed in detail~\cite{Petre:2013:SDA:2535028}. Although the concepts of traffic lights, lanes, and intersections are common and appear to be simple, building a traffic simulator to represent these relationships and events in real time turns out to be challenging.

@@

Figure~\ref{fig:trafficsim} illustrates a GRL diagram from the traffic simulator design exercise. An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system (\texttt{Student}, Figure~\ref{fig:trafficsim}), or the system itself ( \texttt{Traffic Tycoon}, Figure~\ref{fig:trafficsim}). Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied. Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way. Softgoals (e.g.  \texttt{Realistic simulation)} are often more related to non-functional requirements, whereas goals (such as  \texttt{Generate cars}) are more related to functional requirements. Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals. In Figure~\ref{fig:trafficsim}, some of the tasks are  \texttt{Create new cars} and \texttt{Keep same cars}. In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \texttt{External Library}, Figure~\ref{fig:trafficsim}).

@@

GRL has six evaluation algorithms which are semi-automated and allow the analysis of alternatives and design decisions by calculating the satisfaction value of the intentional elements across multiple diagrams quantitatively, qualitatively or in a hybrid way. The satisfaction values from intentional elements in GRL can also be propagated to use case maps elements.  jUCMNav, GRL tool-support, also allows for adding new GRL evaluation algorithms~\cite{jUCMNav}. GRL also has the capability to be extended through metadata, links, and external OCL constraints. This allows GRL to be used in many domains without the need to change the whole modeling language. This feature also helps us to apply our argumentation to other domain such as compliance, which we explain in more detail in Section~\ref{sect:goalmodeling:openissues}.

@@

These are the type of the questions that we cannot answer just by looking at the GRL models. The model in Figure~\ref{fig:trafficsim} does not contain information about discussions that let up to the resulting elements of the model, such as various clarification steps for the naming, or alternatives that have been considered for the relationships. In this article we aim to address this shortcoming.

@@

So, for example, we can say that 
\begin{itemize}
\item[] We have goal  \texttt{Generate traffic},
\item[]  \texttt{Keep same cars} will realize goal  \texttt{Generate traffic},
\item[] Which will promote the value  \texttt{Simple design}
\item[] \textit{Therefore} 
\item[] We should perform action  \texttt{Keep same cars}
\end{itemize}

@@

\begin{enumerate}
\item Will the action bring about the desired goal?
%\item Does the goal promote the value stated?
\item Are there alternative ways of realizing the same goal?
\item Are there alternative ways of promoting the same value?
%\item Does doing the action have a side effect which demotes the value?
\item Does doing the action have a side effect which demotes some other value?
\item Does doing the action promote some other value?
%\item Does doing the action preclude some other action which would promote some other value?
\item Is the action possible?
\item Can the desired goal be realized?
\item Is the value indeed a legitimate value?
\end{enumerate}

These critical questions can point to new arguments that might counter the original argument. Take, for example, critical question 4: if we find that  \texttt{Keep same cars} actually negatively influences the value  \texttt{Realistic simulation}, we have a counterargument to the above argument. Another way to counter an argument for an action is to suggest an alternative action that realizes the same goal (question 2) or an alternative goal that promotes the same value (question 3). For example, we can argue that  \texttt{Create new cars} also realizes the goal  \texttt{Generate traffic}, which gives us a counterargument to the original argument -- to generate traffic by simply keeping the cars that disappear off the screen and have them wrap around to the other side of the screen -- that also follows PRAS.

In argumentation, counterarguments are said to \emph{attack} the original arguments (and sometimes vice versa). In the work of Atkinson et al.~\cite{atkinson2007}, arguments and their attacks are captured as an \emph{argumentation framework} of arguments and attack relations as introduced by Dung~\cite{Dung1995}\footnote{Full definitions of Dung's~\cite{Dung1995} frameworks and semantics will be given in section \ref{sect:gmas}. In this section, we will briefly discuss the intuitions behind these semantics.}. Figure \ref{fig:pras:example} shows an argumentation framework with three arguments from the above example: the argument for  \texttt{Keep same cars} (A1), the argument for \texttt{Create new cars} (A3), and the argument that \texttt{Keep same cars} demotes the value  \texttt{Realistic simulation} (A2). The two alternative PRAS instantiations are A1 and A3. These arguments mutually attack each other, as \texttt{Keep same cars}  and \texttt{Create new cars} are considered to be mutually exclusive. Argument A2 attacks A1, as it points to a negative side-effect of \texttt{Keep same cars}.

@@

>>>>

Most of the examples in this article, as well as the topic of discussion in the transcripts we analyze, come from the traffic simulator design exercise. In this exercise, designers are provided a problem description, requirements, and a description of the desired outcomes. The problem description is given in full in Appendix~\ref{ch:designprompt}, %TODO SG: shown as ??
 and is summarized as follows: The client of the project is Professor E, who teaches civil engineering at UCI. %TODO SG: Make it anonymized. For example use a university in California (if this is UC Irvine)
 It is the task of the designer to specify a system in which the professor can teach students how various theories around traffic lights works, such as queuing theory. To this end, a piece of software has to be developed in which students can create visual maps of an area, regulate traffic, and so forth. The original version of the problem descrption~\cite{UCIworkshop} is well known in the field of design reasoning since it has been used in a workshop\footnote{\url{http://www.ics.uci.edu/design-workshop/}}, and transcripts of this workshop have been analyzed in detail~\cite{Petre:2013:SDA:2535028}. Although the concepts of traffic lights, lanes, and intersections are common and appear to be simple, building a traffic simulator to represent these relationships and events in real time turns out to be challenging.

@@

Figure~\ref{fig:trafficsim} illustrates a GRL diagram from the traffic simulator design exercise. An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system (\texttt{Student}, Figure~\ref{fig:trafficsim}), or the system itself ( \texttt{Traffic Tycoon}, Figure~\ref{fig:trafficsim}). Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied. Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way. Softgoals (e.g.  \texttt{Realistic simulation)} are often related to non-functional requirements, whereas goals (such as  \texttt{Generate cars}) are related to functional requirements. Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals. In Figure~\ref{fig:trafficsim}, some of the tasks are  \texttt{Create new cars} and \texttt{Keep same cars}. In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \texttt{External Library}, Figure~\ref{fig:trafficsim}).

@@

GRL has six evaluation algorithms which are semi-automated and allow the analysis of alternatives and design decisions by calculating the satisfaction value of the intentional elements across multiple diagrams quantitatively, qualitatively or in a hybrid way. The satisfaction values from intentional elements in GRL can also be propagated to use case maps elements.  jUCMNav, GRL tool-support, also allows for adding new GRL evaluation algorithms~\cite{jUCMNav}. GRL also has the capability to be extended through metadata, links, and external OCL constraints. This allows GRL to be used in many domains without the need to change the whole modeling language. This feature also helps us to apply our argumentation to other domains such as compliance, which we explain in more detail in Section~\ref{sect:goalmodeling:openissues}.

@@

These are the type of the questions that we cannot answer just by looking at the GRL models. The model in Figure~\ref{fig:trafficsim} does not contain information about discussions that leads to the resulting elements of the model, such as various clarification steps for the naming, or alternatives that have been considered for the relationships. In this article we aim to address this shortcoming.

@@

So, for example, we can say that 
\begin{itemize}
\item[] We have goal  \texttt{Generate traffic},
\item[]  \texttt{Keep same cars} will realize goal  \texttt{Generate traffic}, %TODO SG: where is this in the model?
\item[] Which will promote the value  \texttt{Simple design} %TODO SG: where is this in the model?
\item[] \textit{Therefore} 
\item[] We should perform action  \texttt{Keep same cars}
\end{itemize}

@@

\begin{enumerate}
\item Will the action bring about the desired goal?
%\item Does the goal promote the value stated?
\item Are there alternative ways of realizing the same goal?
\item Are there alternative ways of promoting the same value?
%\item Does doing the action have a side effect which demotes the value?
\item Does performing the action have a side effect which demotes some other value? %TODO SG: do you mean correlation link?
\item Does performing the action promote some other value?
%\item Does doing the action preclude some other action which would promote some other value?
\item Is the action possible?
\item Can the desired goal be realized?
\item Is the value indeed a legitimate value?
\end{enumerate}

These critical questions can point to new arguments that might counter the original argument. Take, for example, critical question 4: if we find that  \texttt{Keep same cars} actually negatively influences the value  \texttt{Realistic simulation}, we have a counterargument to the above argument.  %TODO SG: be careful with the concepts and semantics of links and elements in GRL. If you talk about side effect this indicates a correlation link. If you talk about impact or influence this indicates contribution link. In CQ4 you say side effect, here you say influence and in the model you show contribution link. Finally which one???? 
%TODO SG: Keep the same car in the model positively impacts the realistic simulation. But you say different here... VERY confusing.
Another way to counter an argument for an action is to suggest an alternative action that realizes the same goal (question 2) or an alternative goal that promotes the same value (question 3). For example, we can argue that  \texttt{Create new cars} also realizes the goal  \texttt{Generate traffic}, %TODO SG: where is this in the model?
which gives us a counterargument to the original argument -- to generate traffic by simply keeping the cars that disappear off the screen and have them wrap around to the other side of the screen -- that also follows PRAS. %TODO SG: what do you mean?

In argumentation, counterarguments are said to \emph{attack} the original arguments (and sometimes vice versa). In the work of Atkinson et al.~\cite{atkinson2007}, arguments and their attacks are captured as an \emph{argumentation framework} of arguments and attack relations as introduced by Dung~\cite{Dung1995}\footnote{Full definitions of Dung's~\cite{Dung1995} frameworks and semantics will be given in section \ref{sect:gmas}. In this section, we will briefly discuss the intuitions behind these semantics.}. Figure \ref{fig:pras:example} shows an argumentation framework with three arguments from the above example: the argument for  \texttt{Keep same cars} (A1), the argument for \texttt{Create new cars} (A3), and the argument that \texttt{Keep same cars} demotes the value  \texttt{Realistic simulation} (A2). The two alternative PRAS instantiations are A1 and A3. These arguments mutually attack each other, as \texttt{Keep same cars}  and \texttt{Create new cars} are considered to be mutually exclusive. Argument A2 attacks A1, as it points to a negative side-effect of \texttt{Keep same cars}. %TODO SG: again negative side-effect!

@@

%TODO SG: VERY IMPORTANT! I am confused with this section. Are you explaining the GRL model and discussing how you are going to argue about it in order to change it? Or are you discussing how the links and elements in the GRL are created! Either of both, you should explain it clearly. You wrote this section and section 4 as if everyone has lots of background and they are in your head. Make them clear... If for example you are arguing that the + contribution link for keept same cars is not correct and you are changing it to negative side-effect, you should explain it. SO SO SO confusing... especially that nothing is matched between the model and the text.
